---
title: "FinalProject_EDA"
author: "Joshua Chadwick"
date: "2024-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
```

# Exploring our Social Media Survey Data
```{r}
data <- read.csv("../inputs/train.csv")
data <- data %>%
  drop_na() # Removing the blank lines in the data.
```

### Transforming our Categorical Data into Numerical Data
----------------
We will be using "One-Hot Encoding" for this transformation which separates categorical data into its own columns and gives a binary classification (yes or no) for each row.

**Mutating the Gender variable**:
```{r}
data <- data %>%
  mutate(dummy = 1) %>%
  pivot_wider(names_from = Gender, values_from = dummy, values_fill = 0) # Fill spaces without the given gender category with 0.
```

**Mutating the Platform variable**:
```{r}
data <- data %>%
  mutate(dummy = 1) %>%
  pivot_wider(names_from = Platform, values_from = dummy, values_fill = 0) # Fill spaces without the given gender category with 0.
```

**Mutating the Dominant_Emotion variable**:
```{r}
data <- data %>%
  mutate(dummy = 1) %>%
  pivot_wider(names_from = Dominant_Emotion, values_from = dummy, values_fill = 0) # Fill spaces without the given gender category with 0.
```

**Remove the User_ID column**:
It's irrelevant to our training.
```{r}
data <- select(data, -User_ID)
```

### Export our cleaned and processed data
```{r}
write.csv(data, "cleaned_train.csv", row.names=FALSE)
```